---
title: "HFI Report"
output: html_document
---

```{r setup, include=FALSE}
source("./scripts/chart_1_pf_identity.R")
source("./scripts/chart_2_world.R")
source("./scripts/chart_3_scatterplot.R")
source("./scripts/table_summary.R")
source("./scripts/Summary.R")
```

## Introduction
Here, we look to investigate what factors shape human freedom across the world. The [Human Freedom Index](https://www.kaggle.com/gsutters/the-human-freedom-index) is a way to quantify human freedom based off of measures such as personal, civil, and economic freedom. 
It uses 76 distinct indicators in the main areas of: Rule of Law, Security and Safety, Movement, Religion, Association, Assembly, and Civil Society, Expression and Information, Identity and Relationships, Size of Government, Legal System and Property Rights, Access to Sound Money, Freedom to Trade Internationally, Regulation of Credit, Labor, and Business.
The data operates under the assumption that human freedom is is defined as the absence of coercive constraint and is ranked on a scale from 1 (least freedom) to 10 (most freedom). Through all the protests and controversy, we want to see how these freedom indexes have changed over time, as well as what countries have been progressive enough for these indexes to increase.

The data comes from the Cato Institute, the Fraser Institute, and the Liberales Institut at the Friedrich Naumann Foundation for Freedom.

## Summary
For year 2017, the country with the highest human freedom index (HFI) was `r hf_highest` and the country with the lowest HFI was `r hf_lowest`. As for the average human freedom across different regions, North America ranks the highest, followed by Western Europe, East Asia, Oceania, Eastern Europe, Latin America & the Caribbean, Caucasus & Central Asia, South Asia, Sub-Saharan Africa, while Middle East & North Africa ranks the lowest. Comparing 
personal freedom with women freedom in different regions, the data indicates higher women freedom in [Caucasus & Central Asia], [Oceania] and [North America] but lower personal freedom in these regions. Besides, economic freedom sometimes is a vital factor of evaluating human freedom in different regions. It is expected that the data implies an almost the same rank of economic freedom as that of human freedom, with North America ranks the highest, followed by Western Europe, East Asia, Oceania, Eastern Europe, Caucasus & Central Asia, South Asia, Latin America & the Caribbean, Middle East & North Africa and Sub-Saharan Africa the lowest.

```{r}
summary_info
```

## Table
Since the main focus of our data is to quantify human freedom, the overall trend human freedom scores across the countries, region, the correlation between personal freedom and economic freedom, women-specific Personal Freedom Score and its indicator scores, all of which make up the majority part of the data may be of some interest that enable us to understand more about such quantification of human freedom.  

First, we want to examine whether there is a trend of human freedom scores by country over time.  
```{r}
head(trend_overall_countries)
```
From this table, we will be able to tell that the Overall human freedom score across the countries is not showing noticeable improvement from 2008 to 2017, some had slightly improvement while some got worse.  

Then looking at specifically in 2017 in more detail, we explore overall human freedom scores across the region  
```{r}
overall_region_2017
```
In terms of the overall human freedom on average across the region in 2017,
North America ranks the highest, followed by Western Europe, East Asia, Oceania, Eastern Europe, Latin America & the Caribbean, Caucasus & Central Asia, South Asia, Sub-Saharan Africa, Middle East & North Africa being the lowest.  

After looking at the overall human freedom score, we take a look of the personal and economic scores. Here, by creating a new column, we will be able to tell which aspect of scores outweigh the other in each country.  
```{r}
head(ef_pf)

per_over_eco

eco_over_per
```
In general, Western Europe & North America countries have higher personal freedom scores, with some of the lowest economic freedom scores while the others are included in both cases.    

Since the data purposely factor out female freedom scores, it can be informative to exmaine. Here, the focus is the most recent year 2017 and to see the difference across region.  
```{r}
women_pf_2017
```
In terms of the safety and security aspect for women across the region in 2017, Sub-Saharan Africa ranks the highest, followed by Latin America & the Caribbean, Eastern Europe Western Europe, South Asia, Middle East & North Africa, East Asia, Caucasus & Central Asia, Oceania and North America being the lowest.  

To understand the female freedom score clearly, it is best to take a closer look on its indicators' score.  
```{r}
avg_women_pf_2017
```
In terms of the Women-Specific Personal Freedom Indicator across the region on average in 2017, Caucasus & Central Asia ranks the highest, followed by East Asia, Eastern Europe, Latin America & the Caribbean, North America, Oceania, South Asia, Western Europe, Middle East & North Africa, and Sub-Saharan Africa being the lowest.  

Above all, every table of interest present a possible direction to do more exploration on this data.

## Visualizations
<br>
```{r chart1, echo=FALSE}
chart1
```

This chart was intended to demonstrate the differences in personal freedoms in establishing one’s own identity, sexual orientation, and to have intimate and familial relationships with others across different
regions.

Each of the **Legend** labels represents a specific index in personal freedom regarding one's identity and relationships. The explanations for these labels are as follows:
<br>
<br>
`pf_identity_divorce` measures whether women and men **1)** have the same legal
rights to initiate divorce and **2)** have the same requirements for divorce.
<br>
`pf_identity_legal` measures the degree to which people are free to legally
change their sex and gender.
<br>
`pf_identity_sex_female` measures the extent to which sexual relationships
between women are legal.
<br>
`pf_identity_sex_male` measures the extent to which sexual relationships
between men are legal.
<br>
<br>
From the chart above, we can observe that regions like Sub-Saharan Africa, South Asia, and Middle East & North Africa, which mainly consist of countries that are considered less "developed", exhibit significantly lower mean index for three or more of the categories considered. Of those areas, the freedom of divorce and the freedom to have male-male relationships have the lowest indexes.
<br>
<br>
This graph demonstrates the differences in personal freedoms across geographic regions.

```{r chart2, echo=FALSE}
map_params <- mapCountryData(joined,
  catMethod = "fixedWidth",
  nameColumnToPlot = "hf_score",
  missingCountryCol = "white",
  addLegend = FALSE,
  mapTitle = "Freedom Index by Country - 2017",
  colourPalette = colour_palette)
do.call(addMapLegend, c(map_params, legendWidth = 0.5, legendMar = 2))
```

This chart was intended to show a geographical representation of the overall human freedom index values across countries. Yellow indicates low freedom and blue indicates a higher level of freedom. The white represents countries without values. From the graphic, it is clear that countries that are generally considered more "developed" (North America, Australia, and Western Europe) also have a higher level of human freedom.
<br>
<br>

```{r chart3, echo=FALSE}
chart3
```

This chart was intended to show the relationship of personal and economic freedom in 2017. They're color coded by region. Middle East & North Africa have the lowest personal and economic freedoms, and Africa as a whole is the region with the most widespread freedoms, ranging anywhere from 2-8. Clumped heavily at the top of the scatter-plot is Western Europe, where personal and economic freedoms are the highest. From this scatterplot, we can see that economic and personal freedom are positively correlated, meaning generally those who have higher economic freedom often have higher personal freedom.
